Kepler's Demographic Cliff: Ages of Young Stars and Planets in the Kepler Field

### 2024-05-15 paper/ms.tex
update ms

However, beyond rotation-based ages, recent years have also yielded
improvements in the lithium age scale.  Lithium ages include two
distinct regimes: depletion boundary ages for M dwarfs and brown dwarfs in star
clusters, and the less precise decline-based ages for individual field
FGK stars \citep{Soderblom_2010}.  Here we focus on the latter
approach, which relies on the observation that the Li abundances of
partially-convective stars gradually decline as they age
\citep[e.g.][]{2005A&A...442..615S}.  The theoretical explanation for
this decline is debated
\citep[e.g.][]{1995ApJ...441..865C,2010ApJ...716.1269D,2019MNRAS.485.4052C}.
Empirical understanding has however improved due to the work by
\citet{Jeffries_2023}, who built a model for how the equivalent width
(EW) of the \ion{Li}{1} 6708\,\AA\ doublet evolves as a function of
stellar effective temperature and age for a set of 6{,}200 stars in 52
open clusters.  Two-sided lithium ages are useful for Kepler (FGK)
stars between $\approx$0.03-0.5\,Gyr, with a strong dependence on
spectral type since the K dwarfs lose their surface lithium much
faster than G dwarfs.  The precision of the ages reported by the
\citet{Jeffries_2023} method in this regime are in the range of
0.3-0.5~dex.

@@

Our default source for stellar temperatures and surface gravities is
the Gaia-Kepler Stellar Properties Catalog (GKSP;
\citealt{Berger_2020a_catalog}).  The GKSP parameters were reported
for stars with ``AAA'' 2MASS photometry, measured parallaxes in Gaia
DR2,  and $g$-band photometry from either the KIC or the Kepler-INT
survey.  The parameters themselves were derived using
\texttt{isoclassify} \citep{2017ApJ...844..102H} to interpolate over
the MIST isochrone grids \citep{Choi_2016,2016ApJS..222....8D}, given
the SDSS $g$ and 2MASS $K_{\rm s}$ photometry, the Gaia DR2
parallaxes, and $V$-band extinction from the
\citet{2018MNRAS.478..651G} reddening map.  The resulting stellar
parameters are available for \fracstarswithprotwithbtwenty\ of the
\nuniqstarsantosrot\ Kepler stars with rotation periods.  For the
remaining \fracstarswithprotwithoutbtwenty\ of stars that lack
temperatures and surface gravities from
\citetalias{Berger_2020a_catalog}, we adopt the values reported by
\citet{Santos_2019} and \citet{Santos_2021}, which for these cases
primarily derive from the \citet{Mathur_2017} DR25 Kepler Stellar
Properties Catalog, and are mostly derived from photometry.  In the
planet sample, \frackoisnofpwithprotwithbtwenty\ of the
non-false-positive KOIs with rotation periods have parameters from
\citet{Berger_2020a_catalog}, and the remainder similarly are drawn
from DR25.

@@

\facilities{
  Gaia \citep{Gaia_DR3_2022},
  Kepler \citep{Borucki10},
  Keck:I (HIRES).
}

@@

>>>>

Beyond rotation-based ages, recent years have also yielded
improvements in the lithium age scale.  Lithium ages include two
distinct regimes: depletion boundary ages for M dwarfs and brown dwarfs in star
clusters, and the less precise decline-based ages for individual field
FGK stars \citep{Soderblom_2010}.  Here we focus on the latter
approach, which relies on the observation that the Li abundances of
partially-convective stars gradually decline as they age
\citep[e.g.][]{2005A&A...442..615S}.  The theoretical explanation for
this decline is debated
\citep[e.g.][]{1995ApJ...441..865C,2010ApJ...716.1269D,2019MNRAS.485.4052C}.
Empirical understanding has however improved due to the work by
\citet{Jeffries_2023}, who modeled the evolution of 
the \ion{Li}{1} 6708\,\AA\ doublet equivalent width (EW) through
time using a set of 6{,}200 stars in 52
open clusters.  Two-sided lithium ages are useful for Kepler (FGK)
stars between $\approx$0.03-0.5\,Gyr, with a strong dependence on
spectral type since the K dwarfs lose their surface lithium much
faster than G dwarfs.  The precision of the ages reported by the
\citet{Jeffries_2023} method in this regime are in the range of
0.3-0.5~dex.

@@

Our default source for stellar temperatures and surface gravities is
the Gaia-Kepler Stellar Properties Catalog (GKSP;
\citealt{Berger_2020a_catalog}).  The GKSP parameters were reported
for stars with ``AAA'' 2MASS photometry, measured parallaxes in Gaia
DR2,  and $g$-band photometry available from either SDSS or else Kepler-INT
\citep{2012AJ....144...24G}.  The parameters themselves were derived using
\texttt{isoclassify} \citep{2017ApJ...844..102H} to interpolate over
the MIST isochrone grids \citep{Choi_2016,2016ApJS..222....8D}, given
the SDSS $g$ and 2MASS $K_{\rm s}$ photometry, the Gaia DR2
parallaxes, and $V$-band extinction from the
\citet{2018MNRAS.478..651G} reddening map.  The resulting stellar
parameters are available for \fracstarswithprotwithbtwenty\ of the
\nuniqstarsantosrot\ Kepler stars with rotation periods.  For the
remaining \fracstarswithprotwithoutbtwenty\ of stars that lack
temperatures and surface gravities from
\citetalias{Berger_2020a_catalog}, we adopt the values reported by
\citet{Santos_2019} and \citet{Santos_2021}, which for these cases
primarily derive from the \citet{Mathur_2017} DR25 Kepler Stellar
Properties Catalog, and are mostly derived from photometry.  In the
planet sample, \frackoisnofpwithprotwithbtwenty\ of the
non-false-positive KOIs with rotation periods have parameters from
\citet{Berger_2020a_catalog}, and the remainder similarly are drawn
from DR25.

@@

\facilities{
  Gaia \citep{Gaia_DR3_2022},
  Kepler \citep{Borucki10},
  Keck:I (HIRES),
	2MASS \citep{Skrutskie06},
	SDSS \citep{2000AJ....120.1579Y}.
}

@@

\section{TODO's}

\begin{enumerate}
  \item Compare Mathur2023 ages against my ages, and (more interestingly) against the
    cluster scale using NGC6811, Theia520, Melange3, and CepHer stars.
    Could compare against asteroseis too.
    %
  \item Intro needs better discussion of Mathur2023 and Lu2021
    %
  \item Consider moving the bit list description to its own appendix.
    %
  \item Rerun gyro calculation with denser grid, fixing quantization
    and extending upper limit to 5gyr.  (Will use updated Teff
    uncertainties)
\end{enumerate}
